---
title: "MOT_Sequence_Target_Density"
author: "Freya Joessel"
editor: source
date: today
toc: true
toc-depth: 5
# toc-expand: true
number-sections: true
highlight-style: pygments
toc-location: left
reference-location: margin
citation-location: margin
format:
  html: 
    code-fold: true
    html-math-method: katex
---


```{r preamble}
#| echo: false
#| include: false

library(dplyr)
library(progress)
library(ggplot2)
library(purrr)
library(Hmisc)
library(TEfits)
library(brms)
# librarbrms# library(parallel)

```



```{r load-data}
#| echo: false
#| cache: true
#| include: false

clc_raw.df <- read.csv(file = "Data/psychopy_raw.csv") %>% 
  select(-X)

clc_part_1_raw.df <- clc_raw.df %>% 
  filter(!is.na(trials_train.thisN))

clc_part_2_raw.df <- read.csv(file = "Data/psychopy_part2_df.csv") %>% 
  select(-X)

```

```{r filter-non-learners}
#| echo: false
#| cache: false
#| include: false

clc_part_1_n_trials.df <- clc_part_1_raw.df %>% 
  count(Participant_ID, name = "n_trials") 

clc_part_1_n_trials.plot <- ggplot() + 
  geom_histogram(data = clc_part_1_n_trials.df,
                 aes(x = n_trials))
clc_part_1_n_trials.plot

clc_part_1_n_trials_violin.plot <- ggplot() + 
  geom_violin(data = clc_part_1_n_trials.df,
              aes(x = 1, y = n_trials)) + 
  geom_point(data = clc_part_1_n_trials.df,
              aes(x = 1, y = n_trials))
clc_part_1_n_trials_violin.plot

clc_part_1_n_trials_boxplot.plot <- ggplot() + 
  geom_boxplot(data = clc_part_1_n_trials.df,
              aes(x = 1, y = n_trials))
  # geom_point(data = clc_part_1_n_trials.df,
  #             aes(x = 1, y = n_trials))
clc_part_1_n_trials_boxplot.plot

clc_part_1_n_trials_log.plot <- ggplot() + 
  geom_histogram(data = clc_part_1_n_trials.df,
                 aes(x = log(n_trials)))
clc_part_1_n_trials_log.plot

# Remove Outliers
n_trials_outliers <- boxplot.stats(clc_part_1_n_trials.df$n_trials)$out

clc_part_1_n_trials_with_learner_status.df <- clc_part_1_n_trials.df %>% 
  mutate(is_learner = ifelse(n_trials %in% n_trials_outliers, 
                             "non-learner", "learner"))

clc_part_1_with_learner_status.df <- clc_part_1_raw.df %>% 
  merge(clc_part_1_n_trials_with_learner_status.df)

clc_part_1_non_learners.df <- clc_part_1_with_learner_status.df %>% 
  filter(is_learner == "non-learner")

clc_part_1_only_learners.df <- clc_part_1_with_learner_status.df %>% 
  filter(is_learner == "learner")

print(unique(clc_part_1_non_learners.df$Participant_ID))

```

```{r individual-learning-curves-part-1}
#| echo: false
#| cache: false
#| include: false

colors_4_exemplars <- c('purple', 'violet', 'orange', 'gold')

learning_curves_part_1.plot <- ggplot() + 
  geom_line(data = clc_part_1_only_learners.df, 
            aes(x = trials_train.thisTrialN, 
                y = train_response_distance_idx, 
                color = color_name)) + 
  scale_color_manual(values = colors_4_exemplars) + 
  facet_wrap(facets = "Participant_ID", scales = "free_x")
learning_curves_part_1.plot

```

```{r group-level-learning-curves-part-1}
#| echo: false
#| cache: false
#| include: false

clc_part_1_per_group.df <- clc_part_1_only_learners.df %>% 
  group_by(condition_str, trials_train.thisTrialN) %>% 
  summarize(mean_error_idx = mean(train_response_distance_idx))

clc_part_1_per_group_truncated.df <- clc_part_1_per_group.df %>% 
  filter(trials_train.thisTrialN < 200)

learning_curves_per_group_part_1.plot <- ggplot() + 
  geom_point(data = clc_part_1_per_group.df, 
            aes(x = trials_train.thisTrialN, 
                y = mean_error_idx, 
                color = condition_str), size = 0.5) + 
  geom_smooth(data = clc_part_1_per_group.df, 
              aes(x = trials_train.thisTrialN, 
                  y = mean_error_idx, 
                  color = condition_str)) + 
  scale_color_manual(values = c("#0C7BDC", "#FFC20A")) 
learning_curves_per_group_part_1.plot

learning_curves_per_group_part_1_truncated.plot <- ggplot() + 
  geom_point(data = clc_part_1_per_group_truncated.df, 
            aes(x = trials_train.thisTrialN, 
                y = mean_error_idx, 
                color = condition_str), size = 0.5) + 
  geom_smooth(data = clc_part_1_per_group_truncated.df, 
              aes(x = trials_train.thisTrialN, 
                  y = mean_error_idx, 
                  color = condition_str)) + 
  scale_color_manual(values = c("#0C7BDC", "#FFC20A")) 
learning_curves_per_group_part_1_truncated.plot

```

```{r group-level-learning-curves-by-color-part-1}
#| echo: false
#| cache: false
#| include: false

clc_part_1_per_group.df <- clc_part_1_only_learners.df %>% 
  group_by(condition_str, trials_train.thisTrialN, color_name) %>% 
  summarize(mean_error_idx = mean(train_response_distance_idx))

clc_part_1_per_group_truncated.df <- clc_part_1_per_group.df %>% 
  filter(trials_train.thisTrialN < 200)

learning_curves_per_group_part_1.plot <- ggplot() + 
  geom_point(data = clc_part_1_per_group.df, 
            aes(x = trials_train.thisTrialN, 
                y = mean_error_idx, 
                color = condition_str), size = 0.5) + 
  geom_smooth(data = clc_part_1_per_group.df, 
              aes(x = trials_train.thisTrialN, 
                  y = mean_error_idx, 
                  color = condition_str)) + 
  facet_wrap(. ~ color_name) + 
  scale_color_manual(values = c("#0C7BDC", "#FFC20A")) 
learning_curves_per_group_part_1.plot

```


# Number of trials to criterion by color and condition 

```{r n-trials-to-criterion}
#| echo: false
#| cache: false
#| include: false

count_and_threshold <- function(arr, n, k) {
  # Initialize the output vectors
  count_true <- rep(NA, length(arr))
  threshold_array <- rep(NA, length(arr))

  # Loop through each element
  for (i in 1:length(arr)) {
    # Define the window: from max(1, i - n + 1) to i
    start_idx <- max(1, i - n + 1)
    end_idx <- i
    window <- arr[start_idx:end_idx]

    # Count the number of TRUEs in the window
    count_true[i] <- sum(window)

    # Determine the value for the threshold array
    if (i < n) {
      threshold_array[i] <- -1
    } else {
      threshold_array[i] <- ifelse(count_true[i] == k, 1, 0)
    }
  }

  # Return both results as a list
  list(count_true = count_true, threshold_array = threshold_array)
}

n_trials_window <- 5
n_correct_within_window <- 4

clc_part_1_with_learning_criterion.df <- clc_part_1_only_learners.df %>% 
  select(Participant_ID, condition_str, color_name, trials_train.thisTrialN, 
         train_response_correct, train_response_distance_idx, train_response_distance_dE) %>% 
  group_by(Participant_ID, color_name) %>% 
  mutate(n_correct_in_last_five_trials =
           count_and_threshold(train_response_correct, n_trials_window, n_correct_within_window)$count_true, 
         stop_criterion_fulfilled = 
           count_and_threshold(train_response_correct, n_trials_window, n_correct_within_window)$threshold_array)
  
clc_part_1_n_trials_to_learn.df <- clc_part_1_with_learning_criterion.df %>% 
  group_by(Participant_ID, color_name, condition_str) %>% 
  summarise(n_trials_to_criterion = which(stop_criterion_fulfilled == 1)[1], .groups = "keep")

clc_part_1_n_trials_to_learn_exclude.df <- clc_part_1_n_trials_to_learn.df %>% 
  filter(is.na(n_trials_to_criterion))
clc_part_1_n_trials_to_learn_exclude <- 
  unique(clc_part_1_n_trials_to_learn_exclude.df$Participant_ID)

clc_part_1_n_trials_to_learn_no_na.df <- clc_part_1_n_trials_to_learn.df %>% 
  filter(!(Participant_ID %in% clc_part_1_n_trials_to_learn_exclude)) %>% 
  mutate(color_name = factor(color_name, levels = c("violet", "purple", "orange", "yellow")))

clc_part_1_n_trials_to_learn_no_na_stats.df <- clc_part_1_n_trials_to_learn_no_na.df %>% 
  group_by(color_name, condition_str) %>% 
  dplyr::reframe(n_trials_to_criterion_mean = mean(n_trials_to_criterion), 
                 n_trials_to_criterion_ci95_lower = smean.cl.boot(n_trials_to_criterion)["Lower"], 
                 n_trials_to_criterion_ci95_upper = smean.cl.boot(n_trials_to_criterion)["Upper"], 
                 n_trials_to_criterion_sd = sd(n_trials_to_criterion))

clc_part_1_n_trials_to_learn.plot <- ggplot() + 
  geom_violin(data = clc_part_1_n_trials_to_learn_no_na.df, 
             aes(x = condition_str, y = n_trials_to_criterion, color = color_name)) + 
  geom_jitter(data = clc_part_1_n_trials_to_learn_no_na.df, 
             aes(x = condition_str, y = n_trials_to_criterion, color = color_name),
             width = 0.1) +
  geom_errorbar(data = clc_part_1_n_trials_to_learn_no_na_stats.df, 
                aes(x = condition_str, 
                    ymin = n_trials_to_criterion_ci95_lower, 
                    ymax = n_trials_to_criterion_ci95_upper), 
                width = 0.25, color = "black") +
  geom_point(data = clc_part_1_n_trials_to_learn_no_na_stats.df, 
             aes(x = condition_str, y = n_trials_to_criterion_mean), 
             shape = 4, color = "black", size = 2) + 
  facet_grid(. ~ color_name) +
  scale_color_manual(values = colors_4_exemplars)
clc_part_1_n_trials_to_learn.plot

```

Hypotheses : 
- Start closer to where colors are in the Dark Far condition 
-> On trial 1 for each color 
-> OR look at first block of 8 and second block of 8 because that's when they will have seen all the colors to determine what is light and dark
- Learn faster in the Dark Far condition
-> Use TEFits, fix asymptote at a criterion for "perfect", one rate and one start for each participant
-> We'll know that we're failing to capture some variance linked to the difficulty of the color
-> EXPLORATORY -> See if models converge when we fit a different model for each color (colors are independent, nested within participants)
- Specify two models with wieghts shared between them and find the weights thaht best fit the data
-> Make the generalization plots for all participants at first,
-> We at least want to characterize it before more analyses
-> 


# Position of the clicks in the first 8 trials vs 8 following trials

```{r click-positions-first-eight-vs-next-eight}
#| echo: false
#| cache: false
#| include: false

colors_4_exemplars_data <- 
  unique(clc_part_1_only_learners.df %>% 
           select(color_name, color_position, color)) %>% 
  rename(color_hex = color) %>% 
  arrange(color_position) %>% 
  mutate(color_name = factor(color_name, levels = color_name))

colors_4_exemplars_data_by_condition <- colors_4_exemplars_data %>% 
  mutate(color_position = 255 - color_position, 
         condition_str = "Dark Far") %>% 
  bind_rows(colors_4_exemplars_data %>% mutate(condition_str = "Light Far")) %>% 
  merge(clc_part_1_first_v_next_eight.df %>% select(trials_train.thisTrialN, condition_str, block, color_name))

clc_part_1_first_v_next_eight.df <- clc_part_1_only_learners.df %>% 
  filter(trials_train.thisN < 16) %>% 
  mutate(block = ifelse(trials_train.thisN < 8, "first_eight", "next_eight")) %>%
  mutate(color_name = factor(color_name, levels = colors_4_exemplars_data$color_name))

clc_part_1_first_v_next_eight.plot <- ggplot() + 
  geom_violin(data = clc_part_1_first_v_next_eight.df, 
             aes(x = condition_str, y = train_response_color_index, color = color_name)) + 
  geom_jitter(data = clc_part_1_first_v_next_eight.df, 
             aes(x = condition_str, y = train_response_color_index, color = color_name),
             width = 0.1) +
  geom_hline(data = colors_4_exemplars_data_by_condition,
             aes(yintercept = color_position, linetype = condition_str)) +
  facet_grid(block ~ (trials_train.thisTrialN %% 8)) +
  scale_color_manual(values = colors_4_exemplars_data$color_hex)
clc_part_1_first_v_next_eight.plot

```

# Learning rate and starting accuracy 

- Learn faster in the Dark Far condition
-> Use TEFits, fix asymptote at a criterion for "perfect", one rate and one start for each participant
-> We know that we're failing to capture some variance linked to the difficulty of the color
-> EXPLORATORY -> See if models converge when we fit a different model for each color (colors are independent, nested within participants)

```{r TEfits-dataset}
#| echo: false
#| cache: false
#| include: false

clc_part_1_forTEFits.df <- clc_part_1_only_learners.df %>% 
  select(Participant_ID, condition_str, sequence_type, color_name, 
         trials_train.thisTrialN, signed_distance_idx, train_response_correct, 
         train_response_distance_dE, train_response_distance_idx, cohort)

write.csv(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.csv"))
saveRDS(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.rds"))

```

```{r estimate-asymptotic-performance}
#| echo: false
#| cache: false
#| include: false

clc_part_1_forTEFits_asympt_perf.df <- clc_part_1_forTEFits.df %>% 
  group_by(trials_train.thisTrialN) %>% 
  mutate(n_participants_at_trial = n()) %>%
  ungroup() %>% 
  filter(n_participants_at_trial >= 10) %>% 
  group_by(trials_train.thisTrialN) %>% 
  dplyr::summarize(train_response_distance_idx_med = 
                     median(train_response_distance_idx), 
                   .groups = "drop")

clc_part_1_forTEFits_asympt_perf <- clc_part_1_forTEFits_asympt_perf.df %>% 
  tail(n=50) %>% 
  dplyr::summarize(aymptotic_perf = median(train_response_distance_idx_med)) %>% 
  pull(aymptotic_perf)

```



