---
title: "Color Learning"
author: "Freya Joessel"
editor: source
date: today
toc: true
toc-depth: 5
# toc-expand: true
number-sections: true
highlight-style: pygments
toc-location: left
reference-location: margin
citation-location: margin
format:
  html: 
    code-fold: true
    html-math-method: katex
---
  
  
```{r preamble}
#| echo: false
#| include: false

library(dplyr)
library(progress)
library(ggplot2)
library(purrr)
library(Hmisc)
library(TEfits)
library(brms)
# librarbrms# library(parallel)

```

# Learning rate and starting accuracy 

- Learn faster in the Dark Far condition
-> Use TEFits, fix asymptote at a criterion for "perfect", one rate and one start for each participant
-> We know that we're failing to capture some variance linked to the difficulty of the color
-> EXPLORATORY -> See if models converge when we fit a different model for each color (colors are independent, nested within participants)

```{r TEfits-dataset}
#| echo: false
#| cache: false
#| include: false


if(file.exists(file.path("Data", "clc_part_1_forTEFits.rds"))) {
  clc_part_1_forTEFits.df <- readRDS(file.path("Data", "clc_part_1_forTEFits.rds")) 
} else {
  clc_part_1_forTEFits.df <- clc_part_1_only_learners.df %>% 
    select(Participant_ID, condition_str, sequence_type, color_name, 
           trials_train.thisTrialN, signed_distance_idx, train_response_correct, 
           train_response_distance_dE, train_response_distance_idx, cohort)
  
  write.csv(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.csv"))
  saveRDS(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.rds"))  
}


```

```{r estimate-asymptotic-performance}
#| echo: false
#| cache: false
#| include: false

clc_part_1_forTEFits_asympt_perf.df <- clc_part_1_forTEFits.df %>% 
  group_by(trials_train.thisTrialN) %>% 
  mutate(n_participants_at_trial = n()) %>%
  ungroup() %>% 
  filter(n_participants_at_trial >= 10) %>% 
  group_by(trials_train.thisTrialN) %>% 
  dplyr::summarize(train_response_distance_idx_med = 
                     median(train_response_distance_idx), 
                   .groups = "drop")

clc_part_1_forTEFits_asympt_perf <- clc_part_1_forTEFits_asympt_perf.df %>% 
  tail(n=50) %>% 
  dplyr::summarize(aymptotic_perf = median(train_response_distance_idx_med)) %>% 
  pull(aymptotic_perf)

```


```{r TEfits-exp-fit}
#| echo: false
#| cache: false
#| include: false

if(file.exists(file.path('Data', 'tefits_exp_model.rds'))) {
  model_exp <- readRDS(file = file.path('Data', 'tefits_exp_model.rds'))
} else {
  # Simulation parameters
  n_chains <- 4
  n_cores <- 4
  n_iter <- 10000
  
  # Exponential
  model_exp <- TEbrm(train_response_distance_idx ~ 
                  tef_change_expo3('trials_train.thisTrialN',
                                   parForm = ~ (1 | Participant_ID), 
                                   startForm = ~ condition_str + (1 | Participant_ID),
                                   rateForm = ~ condition_str + (1 | Participant_ID),
                                   asymForm = clc_part_1_forTEFits_asympt_perf),
    # the following are all arguments directly to the `brm()` function call:
    data = clc_part_1_forTEFits.df,
    family = exgaussian(),
    chains = n_chains, # At least 2 chains would be necessary for trustworthy inferences
    cores = n_cores, 
    iter = n_iter, 
    refresh = 10,
    control = list(adapt_delta = .9) # this is just to make things run faster, but will lead to worse convergence
  )
  
  model_exp$participant_coefficients <- data.frame(coef(model_exp)[["Participant_ID"]][,'Estimate',])
  
  model_exp$pred_err <- fitted(model_exp)
  
  saveRDS(object = model_exp, file = file.path('Data', 'tefits_exp_model.rds'))
}

```

```{r TEfits-power-fit}
#| echo: false
#| cache: false
#| include: false

if(file.exists(file.path('Data', 'tefits_pow_model.rds'))) {
  model_pow <- readRDS(file = file.path('Data', 'tefits_pow_model.rds'))
} else {
  
  # Simulation parameters
  n_chains <- 4
  n_cores <- 4
  n_iter <- 10000
  
  # Power
  model_pow <- TEbrm(train_response_distance_idx ~ 
                  tef_change_power3('trials_train.thisTrialN',
                                    parForm = ~ (1 | Participant_ID), 
                                    startForm = ~ condition_str + (1 | Participant_ID),
                                    rateForm = ~ condition_str + (1 | Participant_ID),
                                    asymForm = clc_part_1_forTEFits_asympt_perf),
    # the following are all arguments directly to the `brm()` function call:
    data = clc_part_1_forTEFits.df,
    family = exgaussian(),
    chains = n_chains, # At least 2 chains would be necessary for trustworthy inferences
    cores = n_cores, 
    iter = n_iter,
    control = list(adapt_delta = .9) # this is just to make things run faster, but will lead to worse convergence
  )
  
  # mod1 <- add_criterion(mod1,'loo',pointwise=T)
  
  model_pow$participant_coefficients <- data.frame(coef(model_pow)[["Participant_ID"]][,'Estimate',])
  
  model_pow$pred_err <- fitted(model_pow) # This might take a while to run
  
  saveRDS(object = model_pow, file = file.path('Data', 'tefits_pow_model.rds'))
}

```

```{r TEfits-dataset-plots}
#| echo: false
#| cache: false
#| include: false

model_data_for_plot.df <- data.frame(
  participant_id = clc_part_1_forTEFits.df$Participant_ID,
  train_response_distance_idx = clc_part_1_forTEFits.df$train_response_distance_idx,
  trial_num = model_exp$data$trials_train.thisTrialN,
  exp_estimate = model_exp$pred_err[,'Estimate'],
  exp_q25 = model_exp$pred_err[,'Q2.5'],
  exp_q975 = model_exp$pred_err[,'Q97.5']
  # pow_estimate = model_pow$pred_err[,'Estimate'],
  # pow_q25 = model_pow$pred_err[,'Q2.5'],
  # pow_q975 = model_pow$pred_err[,'Q97.5']
  )

graph1 <- ggplot() + 
   geom_point(data = model_data_for_plot.df,
       aes(x = trial_num,
           y = train_response_distance_idx), 
       size = 0.5) +  
  geom_line(data = model_data_for_plot.df,
       aes(x = trial_num,
           y = exp_estimate, 
           color = participant_id), 
       show.legend = FALSE) + 
  theme_bw() + 
  facet_wrap(. ~ participant_id, scales = "free_x")
graph1
```
