---
title: "Color Learning"
author: "Freya Joessel"
editor: source
date: today
toc: true
toc-depth: 5
# toc-expand: true
number-sections: true
highlight-style: pygments
toc-location: left
reference-location: margin
citation-location: margin
format:
  html: 
    code-fold: true
    html-math-method: katex
---
  
  
```{r preamble}
#| echo: false
#| include: false

library(dplyr)
library(progress)
library(ggplot2)
library(purrr)
library(Hmisc)
library(TEfits)
library(brms)
# librarbrms# library(parallel)

```

# Learning rate and starting accuracy 

- Learn faster in the Dark Far condition
-> Use TEFits, fix asymptote at a criterion for "perfect", one rate and one start for each participant
-> We know that we're failing to capture some variance linked to the difficulty of the color
-> EXPLORATORY -> See if models converge when we fit a different model for each color (colors are independent, nested within participants)

```{r TEfits-dataset}
#| echo: false
#| cache: false
#| include: false


if(file.exists(file.path("Data", "clc_part_1_forTEFits.rds"))) {
  clc_part_1_forTEFits.df <- readRDS(file.path("Data", "clc_part_1_forTEFits.rds")) 
} else {
  clc_part_1_forTEFits.df <- clc_part_1_only_learners.df %>% 
    select(Participant_ID, condition_str, sequence_type, color_name, 
           trials_train.thisTrialN, signed_distance_idx, train_response_correct, 
           train_response_distance_dE, train_response_distance_idx, cohort)
  
  write.csv(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.csv"))
  saveRDS(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.rds"))  
}


```

```{r estimate-asymptotic-performance}
#| echo: false
#| cache: false
#| include: false

clc_part_1_forTEFits_asympt_perf.df <- clc_part_1_forTEFits.df %>% 
  group_by(trials_train.thisTrialN) %>% 
  mutate(n_participants_at_trial = n()) %>%
  ungroup() %>% 
  filter(n_participants_at_trial >= 10) %>% 
  group_by(trials_train.thisTrialN) %>% 
  dplyr::summarize(train_response_distance_idx_med = 
                     median(train_response_distance_idx), 
                   .groups = "drop")

clc_part_1_forTEFits_asympt_perf <- clc_part_1_forTEFits_asympt_perf.df %>% 
  tail(n=50) %>% 
  dplyr::summarize(aymptotic_perf = median(train_response_distance_idx_med)) %>% 
  pull(aymptotic_perf)

```

Next steps 
-> Change the family of the distribution to asym_laplace() because the gaussian component of the distribution is too small compared to the exponential component of the distribution which is captured by the long tail 

-> Specify a bounded prior for the rate such we don't get to extreme values
  nlf(ratePar ~ lowerBound + rateRange*inv_logit(rateParLogit) ) 

  Then
  rateParLogit ~ my_normal_rate_lmem

```{r TEfits-exp-fit}
#| echo: false
#| cache: false
#| include: false

fname_exp_model <- 'tefits_exp_model_v2.rds'

if(file.exists(file.path('Data', fname_exp_model))) {
  model_exp <- readRDS(file = file.path('Data', fname_exp_model))
} else {
  # Simulation parameters
  n_chains <- 2
  n_cores <- 2
  n_iter <- 2000
  
  # Exponential
  model_exp <- TEbrm(train_response_distance_idx ~ 
                  tef_change_expo3('trials_train.thisTrialN',
                                   parForm = ~ (1 | Participant_ID), 
                                   startForm = ~ condition_str + (1 | Participant_ID),
                                   rateForm = ~ condition_str + (1 | Participant_ID),
                                   asymForm = clc_part_1_forTEFits_asympt_perf),
    # the following are all arguments directly to the `brm()` function call:
    data = clc_part_1_forTEFits.df,
    family = asym_laplace(),
    chains = n_chains, # At least 2 chains would be necessary for trustworthy inferences
    cores = n_cores, 
    iter = n_iter, 
    refresh = 10,
    control = list(adapt_delta = .1) # this is just to make things run faster, but will lead to worse convergence (.1 to look at the ourput, above .8 for the actual publication)
  )
  
  model_exp$participant_coefficients <- data.frame(coef(model_exp)[["Participant_ID"]][,'Estimate',])
  
  model_exp$pred_err <- fitted(model_exp)
  
  saveRDS(object = model_exp, file = file.path('Data', fname_exp_model))
  saveRDS(object = model_exp, file = file.path('R:/Projects/Freya/MOT_Project/', fname_exp_model))
}

# model_exp_actualMedian <- 
#   update(model_exp, 
#          family = asym_laplace(link_quantile = 'identity'), 
#          prior = prior_summary(model_exp) + prior(constant(.5), class = 'quantile') )

```


```{r TEfits-power-fit-bounded-rate}
#| include: false
#| echo: false

fname_exp_model_bounded_rate <- 'tefits_exp_model_v2_bounded_rate.rds'

if(file.exists(file.path('Data', fname_exp_model))) {
  model_exp <- readRDS(file = file.path('Data', fname_exp_model))
} else {
  # Simulation parameters
  n_chains <- 3
  n_cores <- 3
  n_iter <- 2000
  delta_param <- 0.1 # change to 0.95 for publication
  
  # Model formula
  bForm <- brmsformula(
    performance ~ pAsym + ( pStart-pAsym )*2^( (1 - trialNum) / pRate ) # standard exponential function. the base-2 log could be replaced with base-e or base-10 or whatever, if you want. The `1` in `1-trialNum` should be equal to the smallest value in your time variable
    , pStartLog ~ condition + (1 | subID) # Wilkenson formula for transformed parameter
    , pRateLogis ~ condition + (1 | subID) # Wilkenson formula for transformed parameter
    , pAsymLog ~ condition + (1 | subID) # Wilkenson formula for transformed parameter
    ,nl = T) +
    nlf(pStart ~ pStartLog^2) + # base-2 log transform (ensures positivity; could also do a scaled logistic if you wanted)
    nlf(pRate ~ 5 + (100 - 5) * inv_logit(pRateLogis) ) + # scaled logistic transform (imposes boundaries of 5 and 100)
    nlf(pAsym ~ pAsymLog^2) # base-2 log transform (ensures positivity; could also do a scaled logistic if you wanted)
  
  model_exp_bounded_rate <- brm(
    bForm
    ,dat
    ,family = asym_laplace(link_quantile = 'identity')
    ,prior = 
      prior(constant(.5),class='quantile') + # ensures estimating median
      prior(normal(0,3),class = 'b', nlpar = 'pStartLog') + # places prior density mostly between 2^-6 and 2^6 [fairly wide]; try `hist(rnorm(1E4,X,X)^2)` to visualize
      prior(normal(0,1.5),class = 'b', nlpar = 'pRateLogis') + # places prior density more-or-less uniformly between our bounds; try `hist(X + (X-X) * plogis(rnorm(1E4,X,X)) )` to visualize alongside the boundaries
      prior(normal(0,3),class = 'b', nlpar = 'pAsymLog') # places prior density mostly between 2^-6 and 2^6 [fairly wide] ; try `hist(rnorm(1E4,X,X)^2)` to visualize
    , chains = n_chains, cores = n_cores, iter = n_iter, 
    , control = list(adapt_delta = delta_param), 
    refresh = 20
  )  
  
  by_participant_params <- 
    aggregate(cbind(trialNum,performance) ~ subID + condition
              ,model_exp_bounded_rate$data
              ,median ) ; for(curParam in c('pStart','pRate','pAsym')){
                
                dTmp <- fitted(myModel
                               ,nlpar = curParam
                               ,newdata = by_participant_params)
                by_participant_params[,curParam] <- dTmp[,'Estimate']
                rm(dTmp)
              }  
  
  saveRDS(object = model_exp_bounded_rate, 
          file = file.path('Data', fname_exp_model_bounded_rate))
  saveRDS(object = model_exp_bounded_rate, 
          file = file.path('R:/Projects/Freya/MOT_Project/', fname_exp_model_bounded_rate))
}




```

```{r TEfits-power-fit}
#| echo: false
#| cache: false
#| include: false

fname_pow_model <- 'tefits_pow_model_v2.rds'
if(file.exists(file.path('Data', fname_pow_model))) {
  model_pow <- readRDS(file = file.path('Data', fname_pow_model))
} else {
  
  # Simulation parameters
  n_chains <- 2
  n_cores <- 2
  n_iter <- 2000
  
  # Power
  model_pow <- TEbrm(train_response_distance_idx ~ 
                  tef_change_power3('trials_train.thisTrialN',
                                    parForm = ~ (1 | Participant_ID), 
                                    startForm = ~ condition_str + (1 | Participant_ID),
                                    rateForm = ~ condition_str + (1 | Participant_ID),
                                    asymForm = clc_part_1_forTEFits_asympt_perf),
    # the following are all arguments directly to the `brm()` function call:
    data = clc_part_1_forTEFits.df,
    family = exgaussian(),
    chains = n_chains, # At least 2 chains would be necessary for trustworthy inferences
    cores = n_cores, 
    iter = n_iter,
    control = list(adapt_delta = .1) # this is just to make things run faster, but will lead to worse convergence
  )
  
  # mod1 <- add_criterion(mod1,'loo',pointwise=T)
  
  model_pow$participant_coefficients <- data.frame(coef(model_pow)[["Participant_ID"]][,'Estimate',])
  
  model_pow$pred_err <- fitted(model_pow) # This might take a while to run
  
  saveRDS(object = model_pow, file = file.path('Data', fname_pow_model))
  saveRDS(object = model_exp, file = file.path('R:/Projects/Freya/MOT_Project/', fname_pow_model))
}

```

```{r TEfits-dataset-plots}
#| echo: false
#| cache: false
#| include: false

model_data_for_plot.df <- data.frame(
  participant_id = clc_part_1_forTEFits.df$Participant_ID,
  train_response_distance_idx = clc_part_1_forTEFits.df$train_response_distance_idx,
  trial_num = model_exp$data$trials_train.thisTrialN,
  exp_estimate = model_exp$pred_err[,'Estimate'],
  exp_q25 = model_exp$pred_err[,'Q2.5'],
  exp_q975 = model_exp$pred_err[,'Q97.5'],
  pow_estimate = model_pow$pred_err[,'Estimate'],
  pow_q25 = model_pow$pred_err[,'Q2.5'],
  pow_q975 = model_pow$pred_err[,'Q97.5']
  )

make_tefit_plot_with_raw_data <- function(model_data, model_estimate) {
  tefit_plot <- ggplot() + 
     geom_point(data = model_data,
         aes(x = trial_num,
             y = train_response_distance_idx), 
         size = 0.5) +  
    geom_line(data = model_data,
         aes(x = trial_num,
             y = {{model_estimate}}, 
             color = participant_id), 
         show.legend = FALSE) + 
    theme_bw() + 
    facet_wrap(. ~ participant_id, scales = "free_x")  
  
  return(tefit_plot)
}

exp_model.plot <- 
  make_tefit_plot_with_raw_data(
    model_data = model_data_for_plot.df,
    model_estimate = exp_estimate
    )

power_model.plot <- 
  make_tefit_plot_with_raw_data(
    model_data = model_data_for_plot.df,
    model_estimate = pow_estimate
    )

ggsave(plot = exp_model.plot, 
       "Figures/exp_model_v1.png", 
       width = 20, height = 14)

ggsave(plot = exp_model.plot + xlim(c(0, 200)), 
       "Figures/exp_model_v1_200_trials.png", 
       width = 20, height = 14)

ggsave(plot = power_model.plot, 
       "Figures/pow_model_v1.png", 
       width = 20, height = 14)

ggsave(plot = power_model.plot + xlim(c(0, 200)), 
       "Figures/pow_model_v1_200_trials.png", 
       width = 20, height = 14)

```

```{r exp-fit-plot-display}
#| echo: false
#| fig-cap: "Exponential model fit with raw data for all participants."
#| cap-location: margin
#| out.width: "100%"
#| fig.align: "center"
#| cache: true

exp_model.plot
```

```{r pow-fit-plot-display}
#| echo: false
#| fig-cap: "Power model fit with raw data for all participants."
#| cap-location: margin
#| out.width: "100%"
#| fig.align: "center"
#| cache: true

power_model.plot
```

