---
  title: "MOT_Sequence_Target_Density"
author: "Freya Joessel"
editor: source
date: today
toc: true
toc-depth: 5
# toc-expand: true
number-sections: true
highlight-style: pygments
toc-location: left
reference-location: margin
citation-location: margin
format:
  html: 
  code-fold: true
html-math-method: katex
---
  
  
```{r preamble}
#| echo: false
#| include: false

library(dplyr)
library(progress)
library(ggplot2)
library(purrr)
library(Hmisc)
library(TEfits)
library(brms)
# librarbrms# library(parallel)

```

# Learning rate and starting accuracy 

- Learn faster in the Dark Far condition
-> Use TEFits, fix asymptote at a criterion for "perfect", one rate and one start for each participant
-> We know that we're failing to capture some variance linked to the difficulty of the color
-> EXPLORATORY -> See if models converge when we fit a different model for each color (colors are independent, nested within participants)

```{r TEfits-dataset}
#| echo: false
#| cache: false
#| include: false


if(file.exists(file.path("Data", "clc_part_1_forTEFits.rds"))) {
  clc_part_1_forTEFits.df <- readRDS(file.path("Data", "clc_part_1_forTEFits.rds")) 
} else {
  clc_part_1_forTEFits.df <- clc_part_1_only_learners.df %>% 
    select(Participant_ID, condition_str, sequence_type, color_name, 
           trials_train.thisTrialN, signed_distance_idx, train_response_correct, 
           train_response_distance_dE, train_response_distance_idx, cohort)
  
  write.csv(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.csv"))
  saveRDS(clc_part_1_forTEFits.df, file.path("Data", "clc_part_1_forTEFits.rds"))  
}


```

```{r estimate-asymptotic-performance}
#| echo: false
#| cache: false
#| include: false

clc_part_1_forTEFits_asympt_perf.df <- clc_part_1_forTEFits.df %>% 
  group_by(trials_train.thisTrialN) %>% 
  mutate(n_participants_at_trial = n()) %>%
  ungroup() %>% 
  filter(n_participants_at_trial >= 10) %>% 
  group_by(trials_train.thisTrialN) %>% 
  dplyr::summarize(train_response_distance_idx_med = 
                     median(train_response_distance_idx), 
                   .groups = "drop")

clc_part_1_forTEFits_asympt_perf <- clc_part_1_forTEFits_asympt_perf.df %>% 
  tail(n=50) %>% 
  dplyr::summarize(aymptotic_perf = median(train_response_distance_idx_med)) %>% 
  pull(aymptotic_perf)

```



```{r TEfits-dataset}
#| echo: false
#| cache: false
#| include: false

# Simulation parameters
n_chains <- 4
n_cores <- 4
n_iter <- 10000

mod1 <- TEbrm(train_response_distance_idx ~ 
                tef_change_expo3('trials_train.thisTrialN',
                                 parForm = ~ condition_str + (1 | Participant_ID), 
                                 asymForm = clc_part_1_forTEFits_asympt_perf),
  # the following are all arguments directly to the `brm()` function call:
  data = clc_part_1_forTEFits.df,
  family = exgaussian(),
  chains = n_chains, # At least 2 chains would be necessary for trustworthy inferences
  cores = n_cores, 
  iter = n_iter,
  control = list(adapt_delta = .9) # this is just to make things run faster, but will lead to worse convergence
)

mod1 <- add_criterion(mod1,'loo',pointwise=T)

mod1$participant_coefficients <- data.frame(coef(mod1)[["Participant_ID"]][,'Estimate',])

mod1$pred_err <- fitted(mod1)

mod1_for_plot.df <- data.frame(
  trial_num = mod1$data$trials_train.thisTrialN,
  estimate = mod1$pred_err[,'Estimate'])

saveRDS(object = mod1, file = file.path('Data', 'tefits_exp_model.rds'))

```
